### 2021-03-22 papers/icse21-src/main.tex
Initial commit to papers branch

\documentclass[10pt,conference]{IEEEtran}

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote.
% If that is unneeded, please comment it out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Packages

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{float}
\usepackage{flushend}
%% \usepackage{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros and definitions
\linepenalty=1000
\widowpenalty10000
\clubpenalty10000

\lstset{
  frame=none,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  morecomment=[l]\%,
}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\setlength{\tabcolsep}{0.5em}

\newcommand{\ourtool}{\textsc{Mutagen}\xspace}

%% \captionsetup[figure]{labelfont={bf},name={Table},labelsep=period}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document

\begin{document}

\title{\ourtool: Faster Mutation-Based Random Testing}

\author{
  \IEEEauthorblockN{Agust\'in Mista}
  \IEEEauthorblockA{
    %% \textit{Department of Computer Science} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    \texttt{mista@chalmers.se}
  }
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstract

\begin{abstract}
We present \ourtool, a fully automated mutation-oriented framework for
property-based testing.
%
Our tool uses novel heuristics to improve the performance of the testing loop,
and it is capable of finding complex bugs within seconds.
%
We evaluate \ourtool by generating random WebAssembly programs that we use to
find bugs in a faulty validator.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Keywords

\begin{IEEEkeywords}
random testing, mutation, heuristics
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections

\section{Introduction}

% Randomly generated values
%
Using randomly generated inputs is a popular and powerful approach when it comes
to testing software \cite{duran1984evaluation}.
%
While plenty, the tools designed for this purpose can be divided into two main
categories: those that use an existing corpus of inputs, and those that generate
such inputs from scratch.

% Corpus-based approaches
%
On one hand, corpus-based tools create new inputs by combining and mutating
seeds from their input corpora using several heuristics
\cite{zalewski2014american, swiecki2017honggfuzz}.
%
Being mostly black-box, this approach is often limited by the lack of knowledge
about the structure of its inputs, although there exist efforts to improve this
situation, e.g., by using grammars describing the syntactic structure of the
generated data while combining seeds in order to produce syntactically valid
test cases \cite{miller2007analysis, wang2019superion, holler2012fuzzing}.

% Generational approaches
%
On the other hand, generational approaches can circumvent this limitation by
generating valid data from scratch using specialized random generators
\cite{eddington2011peach, dharma, grieco2016quickfuzz, grieco2017quickfuzz}.
%
However, writing good random generators by hand is a demanding task that
involves several iterations of trial and error and can take hundreds of hours
\cite{lampropoulos2019coverage}.
%
While there exist tools that tackle this problem by automatically synthesizing
(with varying degrees of complexity) random generators directly from the static
information present in the codebase \cite{mista2018branching,
  mista2019generating, duregaard2012feat, lampropoulos2017generating}, such
approaches are unable to derive suitable generators when the target data
involves complex invariants like those required to generate random programs,
e.g., well-scopedness and well-typedness.
%
In such cases, generators obtained by automatic tools are extremely unlikely to
produce random data with enough structure to penetrate deep layers of our
systems before being discarded.
%
% early as completely invalid.

%% % Random property-based testing
%% %
%% Random property-based testing is a powerful technique for finding bugs.
%% %
%% The expected behavior of a system under test is expressed in terms of
%% a collection of properties that can be executed when given certain inputs.
%% %
%% Then, automated tools can be used for repeteadly probing these properties using
%% randomly generated inputs until any of them gets falsified.

%% % QuickCheck and random generators
%% %
%% In the Haskell realm, QuickCheck is the dominant tool of this sort
%% \cite{hall1996type}.
%% %
%% As well as most other property-based frameworks, QuickCheck provides default
%% random generators for the base types, along with with a library of combinators
%% for writing random generators for any other used-defined algebraic data type
%% (ADT) -- a task that lies on the developers' shoulders.

%% % Automated derivation tools
%% %
%% For simple ADTs, writing custom random generators manually is often just a
%% cumbersome task.
%% %
%% However, as the complexity of the input data grows, developing random generators
%% capable of generating \emph{good} random values can be extremely tricky.
%% %
%% Fortunately, there exists tools that alleviate this problem by automatically
%% synthesizing (with varying degrees of complexity) random data generators
%% directly from the static information present in the codebase.
%% %
%% Most of these tools are solely focused on the ADT definition of the data to be
%% generated, which is often enough to ensure that the derived generators always
%% produce finite well-typed values that can be use for finding bugs in practice.

%% % Limitation
%% %
%% Despite these automated efforts, most of these tools are unable to synthesize
%% random generators capable of producing data satisfying complex invariants.
%% %
%% Perhaps the most clear example of this issue arises while trying to generate
%% random code.
%% %
%% In most host programming languages, invariants like well-scopedness and
%% well-typedness are quite difficult to express directly at the type level.
%% %
%% Thus, any automated mechanism that synthesizes random generators by only looking
%% at the type definition of the generated data will fail to provide acceptable
%% results in practice.
%% %
%% In this scenario, developers are often forced to spend countless hours of trial
%% and error while developing their random generators by hand.

% Mutation-based approach
%
%% Considering that interesting test cases are very unlikely to be produced from
%% scratch when using automatically derived generators,
%
%% other random testing frameworks like FuzzChick (implemented for the Coq
%% programming language) instead focus their efforts on those that actually are.

Recently, Lampropoulos et al. introduced FuzzChick
\cite{lampropoulos2019coverage}, a property-based testing framework for the Coq
programming language that borrows ideas from the fuzzing community to generate
highly structured values while using automatically derived generators.
%
Instead of continuously generating random invalid test cases from scratch,
FuzzChick keeps a queue of interesting previously executed test cases that can
be mutated using type-preserving transformations in order to produce new ones.
%
The logic behind this is simple, mutating an interesting test case in a
small way (at the data constructor level) has a much higher probability of
producing a new interesting test case than generating a new one from scratch
using a na\"ive generator.
%
FuzzChick is likely to preserve the semantic structure of the mutated data, as
mutations are applied directly at the data type level -- the random AST in case
of generating code.

In order to work in practice, FuzzChick relies on execution traces to
distinguish which executed test cases were interesting and are therefore worth
mutating.
%
%% For this purpose, FuzzChick captures the execution trace of the system under
%% test and compares it with the ones already executed.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces -- any other test case is simply discarded.
%
%% %
%% This coverage-based guidance resembles that used by popular software fuzzers
%% like AFL or Honggfuzz, and it helps focusing the mutation mechanism towards a
%% wider variety of tests cases.
%
%% Moreover, it avoids putting effort in undesired cases like continuously mutating
%% irrelevant subexpressions of a test case, or triggering cyclic mutations loops.
%
While powerful, the implementation of FuzzChick is relatively simple, leaving
the door open for future extensions.

In this work we present \ourtool, a random property-based testing framework
implemented in the strongly typed language Haskell that follows the mutational
approach behind FuzzChick, extending it with novel mutation heuristics designed
to converge to counterexamples in fewer tests.
%
We outline these heuristics in the next section and evaluate them in Section
\ref{sec:casestudy}.

%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[t]
  \renewcommand{\figurename}{Table}
  \vspace{-10pt}
  \scriptsize
  \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Bug} &
      \multicolumn{3}{c|}{QuickCheck} &
      \multicolumn{3}{c|}{\ourtool} &
      \multicolumn{3}{c|}{\ourtool (no inheritance)} &
      \multicolumn{3}{c|}{\ourtool (no scheduling)}  \\
      \cline{2-13} &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded \\
      \hline
      1 & N/A & 14.8 & 1000000 & 2.16 & 2194 & 11706.7 & 4.06 & 4079 & 15123.3 & 3659.45 & 101536.4 & 134107.1 \\
      \hline
      2 & N/A & 12.7 & 1000000 & 0.07 & 83.7  & 131.6 & 0.23 & 159.5 & 206.5 & 0.33 & 261.9 & 234 \\
      \hline
      3 & N/A & 14.9 & 1000000 & 0.03 & 56.7 & 91.8 & 0.04 & 51 & 71.9 & 0.28 & 217.2 & 107.1 \\
      \hline
      4 & N/A & 13.7 & 1000000 & 0.21 & 157.2 & 314.2 & 0.63 & 427.2 & 903.3 & 10.85 & 6740.6 & 7372.2 \\
      \hline
      5 & N/A & 14.3 & 1000000 & 6.45 & 6464 & 34856 & 0.82 & 560.1  & 1543.6 & 454.26 & 31577.9 & 43837.5 \\
      \hline
    \end{tabular}
    \vspace{-2pt}
    \caption{\label{fig:results} Time to first failure, passed and discarded
      tests accross different bugs for the faulty WebAssembly validator. Mean values computed after 10 executions.}
  \vspace{-15pt}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%

\section{\ourtool's Heuristics}
\label{sec:heuristics}

\noindent This section introduces some of \ourtool's novel heuristics.

\paragraph{Exhaustive Uniform Mutations}

Perhaps the biggest distinction between FuzzChick and \ourtool lies in how
mutations are applied over test cases.
%
Along with random data generators, FuzzChick automatically derives simple
type-preserving mutators.
%
These mutators work in a simple recursive manner: with uniform probability, they
either mutate a node at the top level of the given value, or apply a mutation
recursively to one of its subexpressions.
%
This mechanism favors mutations to happen on the shallower levels of their
inputs, while deeper mutations are unlikely to happen due to the multiplicative
decline of their probability on each recursive call.

To find complex bugs, we believe that mutations should be able to occur deep
inside of the generated values in a reasonable proportion.
%
%% If our target is to generate data with a complex and deep structure, e.g.,
%% program's ASTs, we certainly want every deep subexpression of a test candidate
%% to be likely to be mutated at some point.
%%
To allow this, the mutators derived by our tool follow a different approach.
%
\emph{They first traverse the input, collecting the path to each mutable
  subexpression.}
%
Then, the testing loop schedules mutations targeted over each one of them,
ensuring that all of them are mutated in a timely manner.

Furthermore, for a given target subexpression, the testing loop produces and
tests every possible mutation exhaustively, reducing the reliance on randomness
in order to find bugs.
%
%% ensuring that if a test candidate is one mutation away
%% from discovering a bug, then the bug will be found deterministically.
%
This approach is inspired by exhaustive testing tools like SmallCheck
\cite{runciman2008smallcheck} or Korat \cite{boyapati2002korat}.
%
However, testing mutations exhaustively comes attached to a high testing cost
per test candidate, so the next heuristic is, in part, focused on reducing the
time complexity of the testing loop of our tool.

\paragraph{Mutation Inheritance}

The second mutation heuristic focuses on preserving the semantically important
subexpressions generated on previous steps.
%
For this purpose, child mutants keep a record of the subexpressions that were
already mutated by their ancestors.
%
This allows each mutant to focus on the previously untouched subexpressions, as
well as the ones freshly generated by its parent.

As a consequence, this heuristic greatly reduces the available positions where
mutations can occur on each new mutant.
%
This, in turn, helps to cope with the cost of running mutations exhaustively, as
described above.

\paragraph{Mutants Scheduling}

FuzzChick keeps a queue of mutation candidates obtained by analyzing the
execution trace of every executed mutant.
%
If a candidate executes a completely new branch in the code, it is inserted at
the end of the queue, and the testing loop will first have to process every
candidate ahead of it before it can start mutating this new (and likely more
interesting) test case.

To account for this, our tool uses a preemptive (FIFO) schedule with priority for
scheduling mutant candidates.
%
In this setting, we capture the depth at which each new execution trace differs
from all the previous ones.
%
This depth is later used to give more priority to those candidates that
``discover'' new parts of the code at earlier stages, favoring a wider
traversal of the execution trace space in less time.

%%%%%%%%%%%%%%%%%%%%

\section{Case Study: WebAssembly Validator}
\label{sec:casestudy}

This section evaluates how the heuristics implemented in our \ourtool affect the
testing performance.
%
For this purpose, we use the validator of the WebAssembly programming language \cite{haas2017bringing}
as a case study.
%
%% WebAssembly is recent a stack-based low level programming language that can run
%% natively on most web browsers, and that is proposed as a compilation target for
%% a variety of higher level languages.
%% %
%% In practice, WebAssembly modules need to be validated before they can be used
%% --- a process similar to type checking in other languages.
%
We took an exiting Haskell implementation of the WebAssembly validator
\cite{wasm-lib} and injected several bugs into it.
%
These bugs let invalid WebAssembly modules pass the validation process.
%
Then, we tested for false positives of the faulty validator by comparing it
against the output of the validator from the WebAssembly reference
implementation (written in OCaml).
%
This was expressed using the following property:

\vspace{-3pt}
\begin{lstlisting}
prop_valid :: WasmModule -> Result
prop_valid wasm = validHs wasm ==> validOCaml wasm
\end{lstlisting}
\vspace{-3pt}

Using {\small\texttt{validHs}} as a precondition, we avoid executing the
reference validator on invalid modules.
%
Instead, they get automatically discarded and the testing loop continues.

Table \ref{fig:results} shows the performance of our tool against QuickCheck
\cite{claessen2011quickcheck} (the reference tool for random testing in Haskell)
across five different bugs, using the same automatically derived random
generator.
%
To evaluate the effectiveness of our inheritance and scheduling heuristics, we
included the results obtained when they are disabled.

Unsurprisingly, QuickCheck fails to find any of the bugs we planted after
generating more than a million values.
%
This is because the automatically derived generator is virtually unable to
produce valid modules on its own.
%
In comparison, \ourtool is able to find all the bugs in a couple of seconds.
%
Guided by the execution traces, the type-preserving mutations that our tool
applies allows us to run a much larger number of valid WebAssembly modules through
the reference validator, which in turn helps finding bugs faster.
%
As an example of the level of invariants required to find bugs, the following
WebAssembly module AST is a counterexample for bug \#1:

\vspace{-3pt}
\begin{lstlisting}
Module {
  types=*'\textcolor{red}{[}'*FuncType{params=[], results=*'\textcolor{blue}{[]}'*}*'\textcolor{red}{]}'*,
  functions=[
    Function{resultType=*'\textcolor{red}{0}'*, localTypes=[], body=[
      I32Const 0,
      If{resultType=*'\textcolor{blue}{[]}'*, then=*'\textcolor{blue}{[]}'*, else=[I32Const 0]}]]},
  tables=[], mems=[], globals=[], elems=[],
  datas=[], start=Nothing, imports=[], exports=[],
}
\end{lstlisting}
\vspace{-3pt}

\noindent This bug causes false validations by not checking that the type of an
\textbf{else} branch in an \textbf{if} expression matches that of the
\textbf{then} one.
%
%% when the type of an \textbf{else}
%% branch of an \textbf{if} expression does not match the \textbf{then} one ---
%% it can be observed in the example above.
%
%% Despite the actual cause for this counterexample to trigger the bug is quite
%% simple (\textbf{if} branches with different types),
%
Despite this, several other things need to be in place for this value to be
considered valid: the type of the generated function must be declared
beforehand, and the function must refer to it by its index in the types list
(marked in red).
%
Moreover, the type of its actual outputs must match the one declared by its type
(marked in blue).
%
Generating valid modules satisfying such invariants is extremely rare using
automatically derived generators.\looseness=-1

If we look at the effect of our heuristics, we can observe that mutation
inheritance can provide substantial speedups in most cases.
%
%% (e.g., bugs \#1, \#3, and \#4).
%
In the case of bug \#5, however, disabling mutation inheritance seems to
help as a shortcut to find counterexamples faster.
%
This suggests that perhaps a hybrid approach would work best in the general case.

Furthermore, our preemptive scheduling heuristic greatly improves the testing
performance in most cases, being orders of magnitude faster in the case of
bugs \#1, \#4, and \#5.

%% At the light of these results,
The heuristics we developed for \ourtool offer a substantial improvement over
the simple (though still powerful) approach taken by FuzzChick.
%
Our current work focuses on gathering more empirical evidence to strengthen this
claim.

%%%%%%%%%%%%%%%%%%%%

\section{Future Work}

%% We presented \ourtool, an automated framework for property-based testing in
%% Haskell focused on generating complex test cases by progressively applying
%% type-preserving mutations over previously generated ones.
%% %
%% Our tool follows a testing approach similar to that of FuzzChick, extending it
%% with heuristics to improve the performance in terms of convergence speed to
%% counterexamples and reproductibility.

%% \ourtool is guided by the execution trace of the mutated test cases, and favours
%% mutating the ones that discover new parts of the code on each step.
%% %
%% Compared to just using automatically derived generators in a purely random
%% fashion, our tool is capable of finding bugs triggered only by inputs satisfying
%% complex invariants that are extremely unlikey to be produced by such generators
%% alone.

%% In the future, we are focused on extending \ourtool with the ability of producing
%% semantic-preserving paired mutations.
%% %
%% This would entail finding related subexpressions within mutated test cases that
%% need to be handled in tandem in order to preserve their semantic correctness.
%% %
%% With this information at hand, we foresee that it could be possible to train our
%% mutation mechanism so these hard-to-find semantic relations are preserved
%% throughout the testing process.

As for future work, we are focused on extending \ourtool with dependent mutations,
where the existence of certain subexpressions would allow or disallow certain
mutations to occur.
%
Concretely, this could be useful to improve the performance of our tool when
used for testing systems expecting programs as inputs.
%
There, for instance, a mutation that introduces a new identifier would allow
subsequent mutations to refer to it, preserving the well-scopedness of the
program.
%% %
%% We foresee that this extension would require external knowdledge from the
%% developer in order to recognize which kinds of subexpression define identifiers,
%% and which other kinds can use them.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography

\bibliography{references}{}
\bibliographystyle{IEEEtran}

\end{document}