### 2021-03-23 papers/oopsla-21/_TZ_34837-obiwan.tex
WIP

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall, anonymous]{acmart}

% ----------------------------------------

%% Packages

\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\pgfplotsset{compat=1.3}

\makeatletter
\patchcmd{\@verbatim}
  {\verbatim@font}
  {\verbatim@font\footnotesize}
  {}{}
\makeatother

\let\oldv\verbatim
\let\oldendv\endverbatim

\def\verbatim{\par\setbox0\vbox\bgroup\oldv}
\def\endverbatim{\oldendv\egroup\fboxsep0pt \noindent\colorbox[gray]{0.95}{\usebox0}\par}

% ----------------------------------------

%% Rights management information. This information is sent to you when you
%% complete the rights form. These commands have SAMPLE values in them; it is
%% your responsibility as an author to replace the commands and values with
%% those provided to you when you complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2021}
\acmYear{2021}
%% \acmDOI{10.1145/1122445.1122456}

%% These commands are for a JOURNAL article.
\acmConference[OOPSLA '21]{OOPSLA '21: ACM SIGPLAN conference on Systems,
  Programming, Languages, and Applications}{October 17--22 2021}{Chicago, IL}

%% \acmJournal{JACM}
%% \acmVolume{37}
%% \acmNumber{4}
%% \acmArticle{111}
%% \acmMonth{8}

%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
\citestyle{acmauthoryear}

% ----------------------------------------
%% Macros

\newcommand{\quickcheck}{\textit{QuickCheck}\xspace}
\newcommand{\quickchick}{\textit{QuickChick}\xspace}
\newcommand{\fuzzchick}{\textit{FuzzChick}\xspace}
\newcommand{\mutagen}{\textsc{Mutagen}\xspace}

% ----------------------------------------

%% end of the preamble, start of the body of the document source.

\begin{document}

\section{\mutagen: Testing Mutants Exhaustively}
\label{sec:mutagen}

In this section we describe the base ideas behind \mutagen, our CGPT tool
written in Haskell.
%
On top of them, Section \ref{sec:heuristics} introduces heuristics that help
finding bugs faster in certain testing scenarios.

In constrast with \fuzzchick, \mutagen does not employ a power schedule to
assign energy to mutable candidates.
%
In turn, it resorts to mutate them in an exhaustive and precise basis, where
%
\begin{inparaenum}
  \item each subexpression of a mutation candidate is associated with a set of
    deterministic mutations, and
  \item for every mutable subexpression, each of these mutations is evaluated
    \emph{exactly once}.
\end{inparaenum}
%
There is a small exception to this rules that we will introduce soon.

This idea is inspired by exhaustive bounded testing tools like SmallCheck (in
Haskell) or Korat (in Java), that produce test cases exhaustively.
%
In simple words, such tools work by enumerating all possible values of the
datatypes used in the testing properties, and then executing them from smaller
to larger until a a bug is found, or a certain size bound is reached.
%
The main problem with this approach is that the space of all possible test cases
often grows exponentially as we increment the size bound, and the user
experiences what it looks like ``hitting a wall'', where no larger test cases
can be evaluated until we exhausted all the immediately smaller ones \cite{feat}.
%
In consequence, such tools can only be applied to relatively simple systems,
where the space of inputs does not grow extremely fast.

Not to be confused by these tools, in \mutagen we do not enumerate all possible
test cases exhaustively.
%
\mutagen uses random generators to find interesting initial seeds, and only then
proceeds to execute all possible mutations.
%
Moreover, these mutants will be mutated further only if they discover new paths
in the target code, so the testing loop automatically prunes the space of test
cases that are worth mutating.

\subsection{Exhaustive Mutations}

To describe \mutagen's testing loop, we first need to introduce the mechanism
used for testing mutations exhaustively.
%
In contrast to \fuzzchick, where mutators are parameterized random generators,
in \mutagen we define mutations as the set of mutants that can be obtained by
altering the input value at the top-level (the root).
%
In Haskell, we define mutations as:

\begin{verbatim}
type Mutation a = a -> [Mutant a]
\end{verbatim}

Where \texttt{Mutant}s come in two flavours, pure and random:

\begin{verbatim}
data Mutant a = PURE a | RAND (Gen a)
\end{verbatim}

Pure mutants are used most of the time, and encode simple deterministic
transformations over the top-level of the input.
%
Like in \fuzzchick, these tranformation can:

\end{document}