### 2021-06-14 papers/oopsla-21/rebuttal.md
Create rebuttal.md

Overview
========

Changes
=======

Detailed comments
=================

Reviewer #1
-----------

How did you decide on what bugs to inject? How concerned are you about their realism? For the IFC example, there was a systematic strategy; what about for WASM?

Replied below under "Reviewer #2".

Reviewer #2
-----------

given the absence of a direct comparison against FuzzChick, how do you make the case that your technique is an improvement over the state of the art?

The IFC stack machine case study shows a direct comparison between FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of mean-time-to-failure, failure rate. We acknowledge that the introduction makes it unclear that we actually evaluate MUTAGEN against FuzzChick in one of the case studies but not in the other. This should be rephrased.

on a related note, Table 3 suggests that Mutagen needs to generate more tests until failure than FuzzChick in at least 7 of the 20 cases. Based on these results, and given the synthetic nature of the bugs, how can you argue that the Mutagen technique is really better?

Table 3 compares results only for the runs in which FuzzChick actually finds the injected bug --- a distinction that clearly benefits FuzzChick, since it shows a low failure rate in this case study, and was only able to find 5 out of 20 bugs with 100% failure rate, whereas MUTAGEN achieves 100% failure rate across all bugs.

Given this observation, comparing only the MTTF across all runs (bug-found or timeout) can skew the representation of the performance of both tools since timed-out executions inflate the overall calculation. This is clear in Fig. 4, where FuzzChick's MTTF clearly jumps over 10^5ms as soon as the failure rate goes below 100%.

Moreover, we observed that FuzzChick tends to find bugs either relatively fast, or not at all within the 1-hour testing budget. In this light, the unstable behavior of FuzzChick could be alleviated by running it several times in a row using much shorter timeout (e.g. 30 seconds). Under this consideration, for the comparison against FuzzChick to be fair, our tool needs to be reliable (always find the bug) without being (possibly orders of magniture) slower on individual runs. Table 3 shows that this in fact the case, where the A12 measure indicates that MUTAGEN is not prone to need many more tests to find a bug that FuzzChick in most of the cases (except only for bugs #15 and #18, which FuzzChick is not capable of finding with 100% success rate).

Extra clarification: we included mean-tests-to-failure in Table 3 merely as an indication of the order of magniture. This metric, hovewer, is quite sensitive to outliers, which are to be expected due to both tools having to sometimes fall back to pure random generation, which normally follows a geometric distribution to find counterexamples. For this reason, we included the A12 measure between the two sets of measures, which is much more robust against outliers and should be considered over simply mean-tests-to-failure when comparing both approaches.

how were bugs introduced in the benchmarks? If you introduced these manually, what steps were taken to mitigate potential bias?

The IFC stack machine case study uses a systematic approach for introducing bugs, removing a single IFC policy check over a single machine instruction at a time. This was originally introduced by Lampropoulos et al. in their original FuzzChick article. We simply translated this full case study (along with the mechanism introducing the bugs) directly into Haskell.

The WebAssembly case study uses manually-inserted bugs. Such bugs are of two different kind:

* Removing an existing integrity check 
  (to weaken the WebAssembly type-system/validator)
 * Replacing the implementation of an operation with a similar compatible one, e.g., (+) -> (-) 
    (to simulate copy-and-paste induced bugs, inspired by one of the real bugs we found on haskell-wasm while developing this case study)

how much time do the techniques require to generate tests?

While we did not gather specific numbers, we don't expect MUTAGEN to be substantially slower than FuzzChick at generating (or mutating) test cases. In both cases, the complexity is at most linear in the size of the generated (or mutated) test case.

Moreover, Haskell's lazy evaluation poses a particular challenge: test cases will not be fully generated/mutated unless the testing property actually fully evaluates them --- test cases are constructed "on the fly" in tandem with the evaluation of the testing property. This is the case with most random testing tools in Haskell, including QuickCheck.

Reviewer #3
-----------

It seems to be the case that Mutagen is effectively a new driver for QuickCheck. Is that right? How does one invoke Mutagen? Can you drop it in to replace QuickCheck? How is the tracing/coverage achieved?

Yes, MUTAGEN aims to be a drop-in replacement for QuickCheck as much as possible. Our tool is invoked in the same way as QuickCheck, using an IO function as a driver that accepts a testing property.

The major difference lies in that MUTAGEN requires some extra constraints to be satisfied, e.g., property inputs should be Mutable (automatically derivable using Template Haskell).

Tracing/coverage is achieved by instrumenting the user code using a source-to-source plugin bundled with MUTAGEN. The plugin can be enabled in a per-module basis to add or remove tracing granularity if needed.

How does Mutagen's memory profile compare to QuickCheck's and FuzzChick's?

 Memory consumption is QuickCheck is almost negligible, as it is completely black-box and "almost" stateless.
FuzzChick needs extra memory for two purposes: the mutation queues (size/length depends on each particularcase study), and the tracing mechanism (a constant bitmap used to keep track of evaluated edges).
Like FuzzChick, MUTAGEN requires extra space for the mutation queues. In contrast, the "path" tracing mechanism used in MUTAGEN is expected to grow continuously as new executions are discovered. In our experiments, we never observed this to be a problem, and multi-day executions never exhausted the 32GB of memory available in our workstation.

If memory consumption becomes a problem for a certain case study, MUTAGEN can also fall to edge-based tracing, using a constant bitmap like AFL/FuzzChick, at the cost of less trace granularity and a less precise FIFO prioritization of test cases (the only metric available to prioritize them becomes the number of discovered edges.)

How much retesting do you do? Do you ever run the exact same test twice?

To the best of our knowledge, there are two situations when retesting can occur:

Randomly generating the same test case twice while on generation mode. This depends on the quality of the random generator used. 
Mutating a test case back to its immediate ancestor. This can happen when a mutation reverts a previous one. In principle, this could only happen once per mutation batch. We initially introduced a rather complex mechanism to avoid this by keeping track of previously mutated subexpressions and making mutants "inherit" this history from their ancestors, however, the added complexity did not show a big improvement, so we decided to keep the mutation mechanism simple.

In WASM, are you only testing single-module programs (p17)?

Yes. Testing multi-module interactions (a.k.a. WebAssembly scripts) using MUTAGEN would be feasible, although we foresee that a more statefull approach for mutating test cases would be helpful for facilitating interaction between mutated WebAssembly modules --- we are currently investigating this posibility.

You fix 25 as the number of mutations in 'no reset' mode. Is that higher or lower than typical?

In our experiments, this depended heavily on the case study. The WebAssembly case study requires almost no resets (since programs with new executions are found continously), whereas in the IFC stack machine case study we typically observed about 6/7 resets before finding a counterexample. Setting the number of random mutations to 25 was an arbitrary decision to evaluate what happened when this parameter is smaller than the one optimized by MUTAGEN.

Table 3 is just about the IFC example, right? What do these bug numbers correspond to? What's special about bugs 15 and 18 that makes FuzzChick so much better than Mutagen?

Yes. As mentioned above, FuzzChick is better at finding bugs #15 and #18 when it actually finds them. In particular, bug #18 is the second hardest one to find, where FuzzChick manages to find a counterexample in 1 out of 4 executions. In this light, our tool trades speed for effectiveness, finding the bugs reliably by using exhaustive mutations at the cost of (potentially) needing longer to find some pathological bugs.